Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove sampleLine diagram and discussion of it, remove discussion of h0=500m

The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.  Once again, the results are similar on all three grids, and results are shown on the BTF and cut cell grids in figures~\ref{fig:gw-theta}a and \ref{fig:gw-theta}c respectively.  However, examining more closely the anomalies in the lee of the mountain for the cut cell grid, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.  This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).

In a further test, the mountain height is doubled from \SI{250}{\meter} to \SI{500}{\meter} with all other parameter values unchanged.  The same spurious anomaly in potential temperature is again present on the cut cell grid but its amplitude increases.  Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).  We find that the decrease in the Exner function of pressure with height is close to linear on all grids.  The potential temperature increases linearly with height on the BTF and SLEVE grids.  On the cut cell grid, the potential temperature anomalies seen in figure~\ref{fig:gw-theta}d appear as a zig--zag in figure~\ref{fig:gw-exner-theta}.

The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.  This is a manifestation of the Lorenz computational mode \citep{arakawa-konor1996,holdaway2013b}.  In models that include moist processes, the Lorenz computational mode can disrupt clouds and generate spurious precipitation \citep{hollingsworth1995}.

To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.  The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground excite the Lorenz computational mode.

@@

\begin{figure}
	\centering
	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
%
	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).  The Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.  The Lorenz computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.   Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.  The thermal profile with a lesser mountain height of \(h_0 = \SI{250}{\meter}\) (not shown) exhibits a computational mode with smaller amplitude.}
	\label{fig:gw-exner-theta}
\end{figure}

>>>>

The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.  Once again, the results are similar on all three grids, and results are shown on the BTF and cut cell grids in figures~\ref{fig:gw-theta}a and \ref{fig:gw-theta}c respectively.  However, examining more closely the anomalies in the lee of the mountain for the cut cell grid, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.  This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).  \TODO{not present?  I think it is, just much smaller amplitude}

To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.

@@
